Co-citations in context

### 2019-06-13 cocit_r.tex
Minor revisions and noting future revisions in blue type

To study the disciplinary composition of references in our custom datasets, we first used the high level WoS classification of five major research areas: life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities. The two largest of these research areas are physical sciences and life sciences and biomedicine, which contribute on average approximately 35.1\% and 62.8\%, respectively, of the references in WoS over the three years of interest. While there is some variability over the years, according to the Uzzi  {\em et al.} model, we would expect close to 35.1\% of the references cited by the publications in any large network  to be drawn from the physical sciences and roughly 62.8\% of the references to be drawn from the life sciences and biomedicine.
Yet the empirical data present a very different  story: 79.6\% of the references cited in physical sciences publications are from the physical sciences and 89.9\% of the references cited in life sciences and biomedicine publications are from the life sciences and biomedicine. In other words, the empirical data shows a strong tendency of publications to cite papers that are in the same major research area rather than in some other research area.  Thus, there is a strong bias towards citations that are {\em intra-network}. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, even though that percentage varied among disciplines in the eight specialties they investigated (from approximately 39\% to 89\% for 2006). Furthermore, these findings argue that a discipline-indifferent random graph model would exhibit misspecification (in deviating substantially from the empirical data), and supports the  concern about definitions of  innovation and conventionality that are based on deviation from expected values.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=0.6\linewidth]{background-effect}     
\caption{\ Intra-network citation shuffling preserves the disciplinary composition of references. Publications of type Article belonging to the three disciplinary networks (Applied Physics, Immunology, and  Metabolism) were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). Citation shuffling was performed using either our algorithm (\emph{repcs})
 or that of Uzzi et al.~(\emph{umsj}). The disciplinary composition of cited references before and after shuffling was measured as frequencies for each of 153 sub-disciplines (from the extended subject classification in WoS) and expressed as a fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of $1$ indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All eight boxplots are generated from 153 observations each. Null values were set to $1$. Note y-axis: $log_2$ scale.} 
\label{fig:be}
\end{figure}

@@

\begin{table}[ht]
\caption{Model misspecification is reduced by constraining substitutions to disciplinary subnetworks. For the set of journal pairs in common between a disciplinary network and the full WoS dataset, Kullback-Leibler (K-L) divergences
between empirical and simulated journal pair frequencies were computed for the years 1985, 1995, and 2005 for the three disciplinary datasets (applied\_physics, immunology, and metabolism) using either the disciplinary network as background or the WoS superset (all\_wos) to generate the null model (Background). K-L divergence was calculated using the R seewave package \citep{seewave2008}.  The reduction in K-L divergence between simulated and observed data  shows that model misspecification is reduced by  restricting the reference substitutions to the disciplinary subnetwork (our model), as compared to treating all substitutions within WoS as equiprobable.
}
\label{tab:kld}
\centering
\begin{tabular}{| ccccc |} 
  \hline
 Disciplinary Network & Year & Background & K-L Divergence & Ratio \\ 
  \hline
 appl\_physics & 1985 & appl\_physics & 1.21 &  \\ 
  & 1985 & all\_wos & 2.37 & 1.96 \\ 
  & 1995 & appl\_physics & 0.86 &  \\ 
  & 1995 & all\_wos & 2.37 & 2.77 \\ 
  & 2005 & appl\_physics & 0.95 &  \\ 
  & 2005 & all\_wos & 2.35 & 2.47 \\ 
    \hline
  immunology & 1985 & immunology & 0.75 &  \\ 
  & 1985 & all\_wos & 1.68 & 2.24 \\ 
   & 1995 & immunology & 0.78 &  \\ 
   & 1995 & all\_wos & 1.70 & 2.19 \\ 
   & 2005 & immunology & 0.73 &  \\ 
  & 2005 & all\_wos & 1.92 & 2.63 \\ 
    \hline
 metabolism & 1985 & metabolism & 1.11 &  \\ 
   & 1985 & all\_wos & 2.24 & 2.02 \\ 
  & 1995 & metabolism & 1.07 &  \\ 
  & 1995 & all\_wos & 2.33 & 2.17 \\ 
     & 2005 & metabolism & 1.19 &  \\ 
   & 2005 & all\_wos & 2.60 & 2.18 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

To study the disciplinary composition of references in our custom datasets, we first used the high level WoS classification of five major research areas: life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities. The two largest of these research areas are \textcolor{blue}{physical sciences and life sciences and biomedicine}, which contribute on average approximately 35.1\% and 62.8\%, respectively, of the references in WoS over the three years of interest. \textcolor{blue}{While there is some variability over the years, according to the Uzzi  {\em et al.} model, we would expect close to 35.1\% of the references cited by the publications in any large network  to be drawn from the physical sciences and roughly 62.8\% of the references to be drawn from the life sciences and biomedicine.}
Yet the empirical data present a very different  story: 79.6\% of the references cited in physical sciences publications are from the physical sciences and 89.9\% of the references cited in life sciences and biomedicine publications are from the life sciences and biomedicine. In other words, the empirical data shows a strong tendency of publications to cite papers that are in the same major research area rather than in some other research area.  Thus, there is a strong bias towards citations that are {\em intra-network}. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, even though that percentage varied among disciplines in the eight specialties they investigated (from approximately 39\% to 89\% for 2006). Furthermore, these findings argue that a discipline-indifferent random graph model would exhibit misspecification (in deviating substantially from the empirical data), and supports the  concern about definitions of  innovation and conventionality that are based on deviation from expected values.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=0.6\linewidth]{background-effect}     
\caption{\ Intra-network citation shuffling preserves the disciplinary composition of references. Publications of type Article belonging to the three disciplinary networks (Applied Physics, Immunology, and  Metabolism) were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (\textcolor{blue}{bg\_local}) or references from all articles in WoS (\textcolor{blue}{bg\_allWoS}). Citation shuffling was performed using either our algorithm (\emph{repcs})
 or that of Uzzi et al.~(\emph{umsj}). The disciplinary composition of cited references before and after shuffling was measured as frequencies for each of 153 sub-disciplines (from the extended subject classification in WoS) and expressed as a fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of $1$ indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All eight boxplots are generated from 153 observations each. Null values were set to $1$. Note y-axis: $log_2$ scale.} 
\label{fig:be}
\end{figure}

@@

\begin{table}[ht]
\caption{Model misspecification is reduced by constraining substitutions to disciplinary subnetworks. For the set of journal pairs in common between a disciplinary network and the full WoS dataset, Kullback-Leibler (K-L) divergences
between empirical and simulated journal pair frequencies were computed for the years 1985, 1995, and 2005 for the three disciplinary datasets (\textcolor{blue}{applied\_physics}, immunology, and metabolism) using either the disciplinary network as background or the WoS superset (all\_wos) to generate the null model (Background). K-L divergence was calculated using the R seewave package \citep{seewave2008}.  The reduction in K-L divergence between simulated and observed data  shows that model misspecification is reduced by  restricting the reference substitutions to the disciplinary subnetwork (our model), as compared to treating all substitutions within WoS as equiprobable.
}
\label{tab:kld}
\centering
\begin{tabular}{| ccccc |} 
  \hline
 Disciplinary Network & Year & Background & K-L Divergence & Ratio \\ 
  \hline
 appl\_physics & 1985 & appl\_physics & 1.21 &  \\ 
  & 1985 & all\_wos & 2.37 & 1.96 \\ 
  & 1995 & appl\_physics & 0.86 &  \\ 
  & 1995 & all\_wos & 2.37 & 2.77 \\ 
  & 2005 & appl\_physics & 0.95 &  \\ 
  & 2005 & all\_wos & 2.35 & 2.47 \\ 
    \hline
  immunology & 1985 & immunology & 0.75 &  \\ 
  & 1985 & all\_wos & 1.68 & 2.24 \\ 
   & 1995 & immunology & 0.78 &  \\ 
   & 1995 & all\_wos & 1.70 & 2.19 \\ 
   & 2005 & immunology & 0.73 &  \\ 
  & 2005 & all\_wos & 1.92 & 2.63 \\ 
    \hline
 metabolism & 1985 & metabolism & 1.11 &  \\ 
   & 1985 & all\_wos & 2.24 & 2.02 \\ 
  & 1995 & metabolism & 1.07 &  \\ 
  & 1995 & all\_wos & 2.33 & 2.17 \\ 
     & 2005 & metabolism & 1.19 &  \\ 
   & 2005 & all\_wos & 2.60 & 2.18 \\ 
   \hline
\end{tabular}
\end{table}
